---
title: "HW3"
output: pdf_document
date: "2024-01-28"
---

## 1

### a

```{r}
library(ggplot2)

generate_matrix <- function(n, p) {
  x <- seq(1, n) / (n + 1)
  X <- outer(x, 0:p, `^`)
  return(X)
}

p <- 1:20
n <- c(30, 50, 100, 200, 500, 1000)
condition_numbers <- data.frame(n = integer(), p = integer(), cond_num = numeric())

for (ni in n) {
  for (pi in p) {
    X <- generate_matrix(ni, pi)
    cond_num <- kappa(X)
    condition_numbers <- rbind(condition_numbers, data.frame(n = ni, p = pi, cond_num = cond_num))
  }
}

ggplot(condition_numbers, aes(x = p, y = cond_num, color = as.factor(n))) +
  geom_line() +
  labs(title = "Condition Numbers Against p",
       x = "p",
       y = "Condition Number",
       color = "n") +
  theme_minimal()

```
As the degree of polynomial p grows, the the condition numbers also grows exponentially, the conditional number grows to around 2e16 when p is 20 and n is 30. The number of samples has a negative impact on the conditional number. As the number of samples get larger, the matrix become more stable. So the polynomial matrix is ill-conditioned and unstable when the degree is large and number of samples is relatively small.

### b

```{r}

generate_matrix <- function(n, p) {
  x <- seq(1, n) / (n + 1)
  X <- poly(x, degree = p)
  return(X)
}

p <- 1:20
n <- c(30, 50, 100, 200, 500, 1000)
condition_numbers <- data.frame(n = integer(), p = integer(), cond_num = numeric())

for (ni in n) {
  for (pi in p) {
    X <- generate_matrix(ni, pi)
    cond_num <- kappa(X)
    condition_numbers <- rbind(condition_numbers, data.frame(n = ni, p = pi, cond_num = cond_num))
  }
}

ggplot(condition_numbers, aes(x = p, y = cond_num, color = as.factor(n))) +
  geom_line() +
  labs(title = "Condition Numbers Against p",
       x = "p",
       y = "Condition Number",
       color = "n") +
  theme_minimal()

```
The conditional number of all the p and n pairs are exactly 1, which indicate that the matrix for orthogonal polynomials are stable.

## 2

### a

```{r}
piecewiseConstant <- function(x, y, L, plot = TRUE){
  intervals <- 2^L
  K = quantile(x, seq(0, 1, len = intervals), type=1)
  pts = rep(0,2*(intervals-1))
  vals <- rep(0,2*(intervals-1))
  
  for (j in 1:intervals) {
    I = (K[j] < x)&(x <= K[j+1])
    if (length(y[I]) > 0 && !all(is.na(y[I]))) {
      fit = lm(y[I] ~ 1)
      pts[2*j-1] = K[j]
      pts[2*j] = K[j+1]
      vals[2*j-1] = coef(fit)
      vals[2*j] = coef(fit)
    }else{
      pts[2*j-1] = K[j]
      pts[2*j] = K[j+1]
      vals[2*j-1] = NA
      vals[2*j] = NA
    }
  }
  
  if(plot){
    plot(x, y, main = "Piecewise Constant Fit", xlab = "x", ylab = "y")
    lines(pts, vals, col="red", lwd = 3)
  }
  return(list(pts = pts, vals = vals))
}
```

### b

```{r}
load("~/workspace/codings/R/04cars.version2.rda")
horsepower <- dat$Horsepower
city_mpg <- dat$City_MPG

horsepower_scaled <- (horsepower - min(horsepower)) / (max(horsepower) - min(horsepower))
L2 <- piecewiseConstant(horsepower_scaled, city_mpg, L = 2, plot = FALSE)
L3 <- piecewiseConstant(horsepower_scaled, city_mpg, L = 3, plot = FALSE)
L4 <- piecewiseConstant(horsepower_scaled, city_mpg, L = 4, plot = FALSE)

plot(horsepower_scaled, city_mpg, xlab = "Horsepower", ylab = "City_Mpg")
lines(L2$pts, L2$vals, col="blue", lwd = 2)
lines(L3$pts, L3$vals, col="green", lwd = 2)
lines(L4$pts, L4$vals, col="red", lwd = 2)
legend("topright", legend = c("L=2", "L=3", "L=4"), col = c("blue", "green", "red"), lty = 1, lwd = 2)
```

