---
title: "HW4"
output: pdf_document
date: "2024-02-06"
---

## 1

### a
```{r}
library(MASS)
data(Boston)

boxplot(medv ~ chas, data = Boston, xlab = "CHAS (0 or 1)", ylab = "MEDV", main = "MEDV by CHAS")

```
The observations with CHAS value 1 have a higher mean MEDV value than observations with CHAS value 0. Observations with CHAS value 0 have outliers.
```{r}
model <- lm(medv ~ chas, data = Boston)
anova(model)

```
The Pr(>F) is less than 0.05 or 0.01. So we reject the null hypothesis which means the mean medv value with chas value 1 and medv with chas value 0  have statistically significant difference. The result consistent with the boxplot, which shows the CHAS value 1 have a significant higher mean MEDV value than observations with CHAS value 0.


### b
```{r}
boxplot(medv ~ rad, data = Boston, xlab = "RAD", ylab = "MEDV", main = "MEDV by RAD")
```
The mean values of MEDV varies a lot with different RAD values.

```{r}
model <- lm(medv ~ rad, data = Boston)
anova(model)
```
The Pr(>F) is 2.2e-16, which is much smaller than 0.05 or 0.01, so we reject the null hypothesis. So the mean medv value are significantly different, which means rad have a significant affect on medv values.

### c

```{r}
library(ggplot2)
Boston$chas_rad <- with(Boston, paste(chas, rad, sep = "-"))
Boston$chas_rad <- as.factor(Boston$chas_rad)

ggplot(Boston, aes(x = chas_rad, y = medv)) +
  geom_boxplot() +
  labs(x = "CHAS and RAD combination", y = "MEDV", title = "MEDV by CHAS and RAD combination") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```
The affect of RAD on MEDV is larger than the affect of CHAS.THe mean values varies a lot within the same CHAS group. The conbination of rad and chas also have an influence on medv values.

```{r}
interaction.plot(Boston$rad, Boston$chas, Boston$medv, 
                 xlab = "RAD", ylab = "MEDV", 
                 legend = TRUE, col = c("red", "blue"), 
                 lty = 1:2, pch = 18:19,
                 main = "Interaction Plot of MEDV by CHAS and RAD")
```
```{r}
model_interaction <- lm(medv ~ chas * rad, data = Boston)
anova(model_interaction)
```
All of chas, rad, and chas:rad have significant affect on medv with alpha be 0.05. It's consistent with the previous F tests. The difference is that the combination of chas:rad also have a significant influence on medv.

### d
```{r}
ggplot(Boston, aes(x = lstat, y = medv, color = factor(chas))) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  labs(color = "CHAS", x = "LSTAT (%)", y = "MEDV", title = "Relationship between LSTAT and MEDV by CHAS") +
  theme_minimal()
```

From the plot we can see that the decrease indeed is effected by the CHAS value.When CHAS is 0, the MEDV decrease faster.

```{r}
model <- lm(medv ~ lstat * chas, data = Boston)
summary(model)
anova(model)
```


From the model fit, we can see that chas has a positive affect on medv and the influence is significant given the F test metric Pr(>F)  is 4.675e-06. 


## 2

### a
```{r}
model_poly <- lm(medv ~ lstat + I(lstat^2) + I(lstat^3), data = Boston)
summary(model_poly)
```
### b
```{r}
library(splines)


lstat_bs <- bs(Boston$lstat, degree = 3,df=4)
model_bs <- lm(medv ~ lstat_bs, data = Boston)

model_ns <- lm(medv ~ ns(lstat, df = 4), data = Boston)

```

### c
```{r}
Boston$poly_pred <- predict(model_poly, newdata = Boston)
Boston$bs_pred <- predict(model_bs, newdata = Boston)
Boston$ns_pred <- predict(model_ns, newdata = Boston)

p <- ggplot(Boston, aes(x = lstat)) +
  geom_point(aes(y = medv, color = "Data Points"), alpha = 0.5) +
  geom_line(aes(y = poly_pred, color = "Polynomial Fit"), size = 1) +
  geom_line(aes(y = bs_pred, color = "B-spline Fit"), size = 1) +
  geom_line(aes(y = ns_pred, color = "Natural Spline Fit"), size = 1)

p <- p + scale_color_manual(values = c("Data Points" = "black", 
                                       "Polynomial Fit" = "red", 
                                       "B-spline Fit" = "green", 
                                       "Natural Spline Fit" = "blue"),
                            name = "Models",
                            labels = c("Data Points", "Polynomial", "B-spline", "Natural Spline"))

p + labs(x = "LSTAT (%)", y = "MEDV") +
  theme_minimal() +
  theme(legend.position = "bottom")


```



